---
title: "Rob Gordon Rebalancing Tool"
output:
  html_document: 
    code_folding: hide
    df_print: kable
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: true 
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: kate
editor_options: 
    chunk_output_type: inline
---

# Setup R Environment
```{r, echo=FALSE}
## Import R Snippets File
knitr::read_chunk('~/Documents/Github/r_snippets/blogdown.R')
```

```{r knitr_options, include=TRUE}

```
```{r}
## Load Required Libraries & Set Paths
library(tidyverse)
library(tidyquant)
library(lubridate)
library(timetk)
library(DT)
library(PerformanceAnalytics)
setwd("~/Documents/GitHub/rob_gordon_2018")
```


# Get Stock Data
```{r}
stocks <- c("SPY", "IYR", "QQQ", "TLT", "GLD", "IWM", "EEM", "DBC", "EFA")
stock_data <- tq_get(stocks
                     , get  = "stock.prices"
                     , from = Sys.Date() - months(12)
                     , to = Sys.Date()) 
stock_data %>% head(10)
```

# Visualize Stock Prices
* Kept only 4 symbols in order to fit
* Will include all in dashboard via interactive filtering
```{r}
# using ggplot2
stock_data %>%
    filter(symbol %in% c("SPY", "GLD", "QQQ", "DBC")) %>%
    ggplot(aes(x = date, y = adjusted, color = symbol)) +
    geom_line(size = 1) +
    geom_bbands(aes(high = high, low = low, close = close), ma_fun = SMA, n = 5,show.legend=TRUE) + 
    labs(title = "Daily Stock Prices",
         x = "", y = "Adjusted Prices", color = "") +
    facet_wrap(~ symbol, ncol = 2, scales = "free") +
    scale_y_continuous(labels = scales::dollar) +
    theme_tq() + 
    scale_color_tq()
```

# Calculate Returns{.tabset .tabset-fade}
## Simple Returns Data
```{r}
mo_returns <- stock_data %>% 
  group_by(symbol) %>%
  tq_transmute(select = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly", 
                 col_rename = "returns") 

mo_returns %>% 
  mutate(returns = round(returns, 6)) %>% 
  arrange(desc(symbol), desc(date)) %>% datatable(class="compact", options = list(scrollX = TRUE, dom = "tp"))

# simple returns = Rt = [ Pt - Pt-1 ]/ Pt-1
# simple_returns <- 
#   stock_data %>% 
#   group_by(symbol, year = year(date)) %>%
#   summarise(adjusted = last(adjusted, order_by = year)) %>% 
#   mutate(return = (adjusted - lag(adjusted, order_by = year)) / lag(adjusted, order_by = year)
#          , return = round(return, 6)) %>% 
#   arrange(desc(symbol), desc(year)) 
#   # %>% filter(symbol == "SPY")
# datatable(simple_returns, class="compact", options = list(scrollX = TRUE, dom = "tp"))
```

## Chart: Returns
* stocks filtered to fit
```{r}
mo_returns %>% 
  filter(symbol %in% c("SPY", "GLD", "QQQ", "DBC")) %>%
    ggplot(aes(x = date, y = returns, fill = symbol)) +
    geom_line(aes(color=symbol)) +
    geom_hline(yintercept = 0, color = palette_light()[[1]]) +
    scale_y_continuous(labels = scales::percent) +
    labs(title = " Monthly Returns",
         subtitle = "stocks limited to fit",
         y = "Monthly Returns", x = "") + 
    facet_wrap(~ symbol, ncol = 2,scales = "free") +
    theme_tq() + 
    scale_fill_tq()
```

## Chart: Smoothed Returns
```{r}
mo_returns %>% 
  filter(symbol %in% c("SPY", "GLD", "QQQ", "DBC")) %>%
    ggplot(aes(x = date, y = returns, fill = symbol)) +
    geom_smooth(aes(color=symbol),method = 'loess' , formula = y ~ x, se=FALSE) +
    geom_hline(yintercept = 0, color = palette_light()[[1]]) +
    scale_y_continuous(labels = scales::percent) +
    labs(title = " Monthly Returns",
         subtitle = "stocks limited to fit",
         y = "Monthly Returns", x = "") + 
    facet_wrap(~ symbol, ncol = 2,scales = "free") +
    theme_tq() + 
    scale_fill_tq()
```

# Growth of $1000 (6mo){.tabset .tabset-fade}
## Data Table
```{r}
init.investment <- 1000
growth <- mo_returns %>% arrange(date) %>%
  mutate(final_value = init.investment * cumprod(1 + returns)) %>%
  arrange(desc(final_value)) 
growth %>% filter(date == max(date)) %>% select(-date)
```

## Visualization
```{r}
growth %>% ggplot(aes(x = date, y = final_value, color = symbol)) +
    geom_line() +
    # geom_smooth(method = "loess") +
    labs(title = "Individual Portfolio: Comparing the Growth of $1K",
         subtitle = "Quickly visualize performance",
         x = "", y = "Investment Value") +
    theme_tq() + theme(legend.position = "right") +
    scale_y_continuous(labels = scales::dollar)
```
## winners (top 5)
```{r}
growth %>% ungroup() %>% filter(date == max(date)) %>% mutate(rank = row_number()) %>% top_n(5,final_value) %>% select(rank, symbol, final_value)
```

# Portfolio Optimization
## Create Time-Series Object for Optimization
```{r}
# convert to ts object
stock_returns <- mo_returns %>% 
  # filter(symbol %in% c("SPY", "GLD", "QQQ", "DBC")) %>%
  spread(key = symbol, value = returns) %>%
  tk_xts(date_var = date) %>% 
  na.omit()
```
## Run Minimum Variance Optimization
* set objective: risk minimization (variance)
* contraint 1: invest all funds
* contrstaint 2: long only
```{r}
# Markowitz Optimization involves minimizing the risk(variance) of returns and maximizing returns.
# https://rdrr.io/cran/PortfolioAnalytics/f/inst/doc/ROI_vignette.pdf
library(PortfolioAnalytics)

# Loading all suggested packages to fix ROI issue
# list.of.packages <- c("foreach", "DEoptim", "iterators", "fGarch", "Rglpk", "quadprog", "ROI", "ROI.plugin.glpk", "ROI.plugin.quadprog", "ROI.plugin.symphony", "pso", "GenSA", "corpcor", "testthat", "nloptr", "MASS", "robustbase")
# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)) install.packages(new.packages)

rm(portf_minvar)
# Create portfolio object
portf_minvar <- portfolio.spec(assets = colnames(stock_returns))
# portf_minvar <- portfolio.spec(assets = colnames(returns.portfolio))
# Add full investment constraint to the portfolio object
portf_minvar <- add.constraint(portfolio=portf_minvar, type="full_investment")
# Add objective to minimize variance
portf_minvar <- add.objective(portfolio=portf_minvar, type="risk", name="var")
# Add long only constraints
portf_minvar <- add.constraint(portfolio=portf_minvar, type="box", min=0, max=1)
 # Run the optimization
opt_gmv <- optimize.portfolio(R=stock_returns, portfolio=portf_minvar, optimize_method="ROI", trace=TRUE)
print(opt_gmv, digits = 5)
#summary(opt_gmv)
print("Optimal Weights") 
wts <- opt_gmv['weights'] %>% as.data.frame() %>% 
  mutate(.,symbol = rownames(.)) %>% select(symbol, weights)
wts
```

# Visualize Portfolio Optimization
```{r, echo=FALSE, warning=FALSE}
weights <- wts$weights %>% as.vector()
symbols <- wts$symbol %>% as.vector()
principal <- 10000
portfolio_growth_monthly <- mo_returns %>%
  # filter(symbol %in% c("DBC", "GLD", "QQQ", "SPY")) %>% 
    tq_portfolio(assets_col   = symbol, 
                 returns_col  = returns, 
                 weights      = weights, 
                 col_rename   = "investment.growth",
                 wealth.index = TRUE) %>%
    mutate(investment.growth = investment.growth * principal)

portfolio_growth_monthly %>%
    ggplot(aes(x = date, y = investment.growth)) +
    geom_line(size = 2, color = palette_light()[[1]]) +
    labs(title = "Optimized Portfolio Growth",
         subtitle = "",
         caption = "Now we can really visualize performance!",
         x = "", y = "Portfolio Value") +
    geom_smooth(method = "loess") +
    theme_tq() +
    scale_color_tq() +
    scale_y_continuous(labels = scales::dollar)

# opt_gmv %>% chart.RiskReward()
# opt_gmv %>% chart.EfficientFrontier(n.portfolios = 1)
# opt_gmv %>% chart.Weights()
# opt_gmv %>% chart.RiskBudget()
```

# BACKTEST
## Get Historical Stock Returns (3 Years)
```{r}
hist_returns <- tq_get(stocks
                     , get  = "stock.prices"
                     , from = Sys.Date() - months(36)
                     , to = Sys.Date()) %>% 
  group_by(symbol) %>%
  tq_transmute(select = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly", 
                 col_rename = "returns") %>% 
  spread(key = symbol, value = returns) %>%
  tk_xts(date_var = date) %>% 
  na.omit()
```

## Perform Backtest
```{r}
# https://rdrr.io/cran/PortfolioAnalytics/f/inst/doc/ROI_vignette.pdf
bt_gmv <- optimize.portfolio.rebalancing(R=hist_returns, portfolio=portf_minvar,
                                         optimize_method="ROI",
                                         rebalance_on="months",
                                         training_period=36-6)

print(bt_gmv) 
```

`r library(dplyr, PerformanceAnalytics)`
# Get Performance {.tabset .tabset-fade}
## descriptive statistics 
`r mo_returns %>% tidyquant::tq_performance(Ra = returns, Rb = NULL, performance_fun = table.Stats)`
## annualized returns
`r mo_returns %>% tidyquant::tq_performance(Ra = returns, Rb = NULL, performance_fun = table.AnnualizedReturns)`
## downside risks
`r mo_returns %>% tidyquant::tq_performance(Ra = returns, Rb = NULL, performance_fun = table.DownsideRisk)`
## downside risk ratios
`r mo_returns %>% tidyquant::tq_performance(Ra = returns, Rb = NULL, performance_fun = table.DownsideRiskRatio)`
## value at risk
`r mo_returns %>% tidyquant::tq_performance(Ra = returns, Rb = NULL, performance_fun = VaR)`
## sharpe ratio
`r mo_returns %>% tidyquant::tq_performance(Ra = returns, Rb = NULL, performance_fun = SharpeRatio)`

# Appendix
## Resources & Notes
* [Performance Analysis Vignette](https://cran.r-project.org/web/packages/tidyquant/vignettes/TQ05-performance-analysis-with-tidyquant.html)
* [Really great article I walked through](https://rstudio-pubs-static.s3.amazonaws.com/333431_8a130f2ff7f84e4faee2a4ac01d6c30e.html)
* [Really great article #2 - same as above or extended?](https://rpubs.com/mr148/333431)
* [Backtesting 4 Portfolio Optimization Strategies In R](https://seekingalpha.com/article/4166115-backtesting-4-portfolio-optimization-strategies-r)
* [Using ROI Solvers for Portfolio Analytics, Vignette](https://rdrr.io/cran/PortfolioAnalytics/f/inst/doc/ROI_vignette.pdf)

## [Example of Efficiency Frontier from Stack Overflow](https://stackoverflow.com/questions/30614457/r-portfolio-analytics-chart-efficientfrontier-function/30650972#30650972)
```{r}
library(PortfolioAnalytics)
  returns <- stock_returns
#  define moment function
  annualized.moments <- function(R, scale=12, portfolio=NULL){
    out <- list()
    out$mu <-    matrix(colMeans(R), ncol=1)
    out$sigma <- cov(R)/scale
    return(out)
  }
# define portfolio
  prt <- portfolio.spec(assets=colnames(stock_returns))
  prt <- add.constraint(portfolio=prt, type="long_only")
  #  leverage defaults to weight_sum = 1 so is equivalent to full_investment constraint
  prt <- add.constraint(portfolio=prt, type="leverage")
  prt <- add.objective(portfolio=prt, type="risk", name="StdDev")
# calculate and plot efficient frontier
  prt_ef <- create.EfficientFrontier(R=12*stock_returns, portfolio=prt, type="mean-StdDev", 
                                      match.col = "StdDev", momentFUN="annualized.moments", scale=12)
  xlim <- range(prt_ef$frontier[,2])*c(1, 1.5)
  ylim <- range(prt_ef$frontier[,1])*c(.80, 1.05)
  chart.EfficientFrontier(prt_ef, match.col="StdDev", chart.assets = FALSE, 
                          labels.assets = FALSE, xlim=xlim, ylim=ylim )
  points(with(annualized.moments(12*stock_returns, scale=12), cbind(sqrt(diag(sigma)), mu)), pch=19 ) 
  text(with(annualized.moments(12*stock_returns, scale=12), cbind(sqrt(diag(sigma)), mu)), 
       labels=colnames(returns), cex=.8, pos=4) 
  chart.EF.Weights(prt_ef, match.col="StdDev")
```

## [Second EF Example from Stack Overflow](https://stackoverflow.com/questions/30614457/r-portfolio-analytics-chart-efficientfrontier-function/30650972#30650972)
```{r}
require(quadprog) 
#min_x(-d^T x + 1/2 b^T D x) r.t A.x>=b
MV_QP<-function(nx, tarRet, Sig=NULL,long_only=FALSE){
  if (is.null(Sig)) Sig=cov(nx)
  dvec=rep(0,ncol(Sig))
  meq=2
  Amat=rbind(rep(1,ncol(Sig)),
             apply(nx,2,mean) )
  bvec=c(1,tarRet )
  if (long_only) {
    meq=1
    Amat=Amat[-1,]
    Amat=rbind(Amat,
               diag(1,ncol(Sig)),
               rep(1,ncol(Sig)),
               rep(-1,ncol(Sig)))
    bvec=bvec[-1]
    bvec=c(bvec,
               rep(0,ncol(Sig)),.98,-1.02)
  }
  sol  <- solve.QP(Dmat=Sig, dvec, t(Amat), bvec, meq=meq)$solution 
}

steps=50
x=returns
 µ.b <- apply(X = x, 2, FUN = mean) 
long_only=TRUE
range.bl <- seq(from = min(µ.b), to = max(µ.b)*ifelse(long_only,1,1.6), length.out = steps) 
risk.bl <- t(sapply(range.bl, function(targetReturn) { 
  w <- MV_QP(x, targetReturn,long_only=long_only) 
  c(sd(x %*% w),w)  }))

weigthsl=round(risk.bl[,-1],4)
colnames(weigthsl)=colnames(x)
weigthsl %>% head()
risk.bl=risk.bl[,1]
rets.bl= weigthsl%*%µ.b
fan=12
plot(x = risk.bl*fan^.5, y = rets.bl*fan,col=2,pch=21,
     xlab = "Annualized Risk ", 
     ylab = "Annualized Return", main = "long only EF with solve.QP")
```

## Example - Multi Layer Optimization
```{r}
# demo(multi_layer_optimization)
```